---
title: "p8105_HW2_al4225.Rmd"
author: "Anjing"
date: "2022-10-03"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




## Problem 2

### Mr. Trash Wheel dataset

Read and clean the Mr. Trash Wheel sheet:

1. specify the sheet in the Excel file and to omit non-data entries (rows with notes / figures; columns containing notes) using arguments in read_excel
2. use reasonable variable names
3. omit rows that do not include dumpster-specific data
4. round the number of sports balls to the nearest integer and converts the result to an integer variable (using as.integer)
```{r}
library(tidyverse)
library(readxl)
Mr_data = read_excel("./data/updated_p2.xlsx",
                     sheet = "Mr. Trash Wheel",
                     range = "A2:N550",
                     col_names = TRUE) %>%
  janitor::clean_names() %>%
  drop_na() %>%
  mutate(sports_balls = as.integer(round(sports_balls, digits = 0)), 
       dumpster = as.numeric(dumpster), 
       year = as.numeric(year), 
       homes_powered = as.integer(round(homes_powered, digits = 0)), 
       name = "mr. trash wheel") %>%
  select(name, everything())

Mr_data
```


### Professor Trash Wheel dataset
Use a similar process to import, clean, and organize the data for Professor Trash Wheel.
```{r}
library(tidyverse)
library(readxl)
Professor_data = read_excel("./data/updated_p2.xlsx",
                     sheet = "Professor Trash Wheel",
                     range = "A2:M97",
                     col_names = TRUE) %>%
  drop_na() %>%
  janitor::clean_names() %>%
  mutate(homes_powered = as.integer(round(homes_powered, digits = 0)), 
       name = "professor trash wheel") %>%
  select(name, everything())

Professor_data
```


### Combine the two datasets. 
To keep track of which Trash Wheel is which, I added an additional variable "name" before combining.

According to the wheel_data dataset, there are `r wheel_data = 
  bind_rows(Mr_data, Professor_data) %>%
  janitor::clean_names() 
nrow(wheel_data) ` observations in wheel_data. Every observation represents a dumpster. There are `r wheel_data = 
  bind_rows(Mr_data, Professor_data) %>%
  janitor::clean_names() 
ncol(wheel_data)` variables in the dataset including the new column "name" I added. Key variables such as the 2 kinds of trash wheels, dumpster number, date, amount of trash they collected in weight and volumes, types of trash such as bottles, polystyrene, cigarette and bags, and the homes they powered are in the dataset. 
```{r}
wheel_data = 
  bind_rows(Mr_data, Professor_data) %>%
  janitor::clean_names() 

nrow(wheel_data) 
ncol(wheel_data)

wheel_data
```

The total number of sports balls collected by Mr. Trash Wheel in 2020 is `r Mr_data %>%
  filter(year == "2020") %>%
  select(sports_balls) %>%
  colSums()`
```{r}
wheel_data %>%
  filter(name == "mr. trash wheel", year == "2020") %>%
  select(sports_balls) %>%
  colSums()
```

The total weight of trash collected by Professor Trash Wheel is `r wheel_data %>%
  filter(name == "professor trash wheel") %>%
  select(weight_tons) %>%
  colSums()`
```{r}
wheel_data %>%
  filter(name == "professor trash wheel") %>%
  select(weight_tons) %>%
  colSums()
```


## Problem 3
### Pols
First, clean the data in pols-month.csv. Use separate() to break up the variable mon into integer variables year, month, and day; replace month number with month name; create a president variable taking values gop and dem, and remove prez_dem and prez_gop; and remove the day variable.

Description: in pols dateset, there are `r pols = read_csv("./data/p3/pols-month.csv",
                col_names = TRUE,
                show_col_types = FALSE) %>%
  janitor::clean_names() %>%
  drop_na() %>%
  separate(mon, into  = c("year", "month", "day"), sep = "-") %>%
  mutate(year = as.integer(year), 
         month = as.integer(month),
         day = as.integer(day), 
         month = month.abb[month]) %>%
  pivot_longer(
    c(prez_gop, prez_dem),
    names_to = "president", 
    values_to = "prez_value"
  ) %>%
  mutate(president = recode(president, "prez_gop" = "gop", "prez_dem" = "dem")) %>%
  filter(!prez_value == 0) %>%
  select(-day, -prez_value) 
nrow(pols)` observations and `r pols = read_csv("./data/p3/pols-month.csv",
                col_names = TRUE,
                show_col_types = FALSE) %>%
  janitor::clean_names() %>%
  drop_na() %>%
  separate(mon, into  = c("year", "month", "day"), sep = "-") %>%
  mutate(year = as.integer(year), 
         month = as.integer(month),
         day = as.integer(day), 
         month = month.abb[month]) %>%
  pivot_longer(
    c(prez_gop, prez_dem),
    names_to = "president", 
    values_to = "prez_value"
  ) %>%
  mutate(president = recode(president, "prez_gop" = "gop", "prez_dem" = "dem")) %>%
  filter(!prez_value == 0) %>%
  select(-day, -prez_value) 
ncol(pols)` variables. It concludes the key variables such as year, month and president. It shows the number of national politicians(governors/senators/representatives) in U.S., and whether the president is democratic or republican every month from Jan/1947 to Jun/2015.
```{r}
library(tidyverse)
pols = read_csv("./data/p3/pols-month.csv",
                col_names = TRUE,
                show_col_types = FALSE) %>%
  janitor::clean_names() %>%
  drop_na() %>%
  separate(mon, into  = c("year", "month", "day"), sep = "-") %>%
  mutate(year = as.integer(year), 
         month = as.integer(month),
         day = as.integer(day), 
         month = month.abb[month]) %>%
  pivot_longer(
    c(prez_gop, prez_dem),
    names_to = "president", 
    values_to = "prez_value"
  ) %>%
  mutate(president = recode(president, "prez_gop" = "gop", "prez_dem" = "dem")) %>%
  filter(!prez_value == 0) %>%
  select(-day, -prez_value)
  
nrow(pols)
ncol(pols)

pols
```

### Snp
Second, clean the data in snp.csv using a similar process to the above. For consistency across datasets, arrange according to year and month, and organize so that year and month are the leading columns.

Description: in snp dateset, there are `r snp = read_csv("./data/p3/snp.csv",
              col_names = TRUE) %>%
  drop_na() %>%
  janitor::clean_names() %>%
  separate(date, into  = c("month", "day", "year")) %>%
  mutate(year = as.integer(year), 
         year = ifelse(year < 23, year + 2000, year + 1900), 
         month = as.integer(month),
         day = as.integer(day), 
         month = month.abb[month]) %>%
  arrange(year, month) %>%
  select(-day) %>%
  select(year, month, everything())
  nrow(snp) ` observations and `r snp = read_csv("./data/p3/snp.csv",
              col_names = TRUE) %>%
  drop_na() %>%
  janitor::clean_names() %>%
  separate(date, into  = c("month", "day", "year")) %>%
  mutate(year = as.integer(year), 
         year = ifelse(year < 23, year + 2000, year + 1900), 
         month = as.integer(month),
         day = as.integer(day), 
         month = month.abb[month]) %>%
  arrange(year, month) %>%
  select(-day) %>%
  select(year, month, everything())
  ncol(snp) ` variables. It concludes the key variables such as year, month and close. It shows the date and the closing values of the S&P stock index from Apr/1950 to May/2015.
```{r}
library(tidyverse)
snp = read_csv("./data/p3/snp.csv",
              col_names = TRUE) %>%
  drop_na() %>%
  janitor::clean_names() %>%
  separate(date, into  = c("month", "day", "year")) %>%
  mutate(year = as.integer(year), 
         year = ifelse(year < 23, year + 2000, year + 1900), 
         month = as.integer(month),
         day = as.integer(day), 
         month = month.abb[month]) %>%
  arrange(year, month) %>%
  select(-day) %>%
  select(year, month, everything())

snp
```

### Unemployment
Third, tidy the unemployment data so that it can be merged with the previous datasets. This process will involve switching from “wide” to “long” format; ensuring that key variables have the same name; and ensuring that key variables take the same values.

Description: in unemployment dateset, there are `r unemployment = read_csv("./data/p3/unemployment.csv",
              col_names = TRUE,
              show_col_types = FALSE) %>%
  drop_na() %>%
  pivot_longer(
    Jan:Dec, 
    names_to = "month", 
    values_to = "unemployment_percent") %>%
  janitor::clean_names() %>%
  arrange(year, month)  
  nrow(unemployment) ` observations and `r unemployment = read_csv("./data/p3/unemployment.csv",
              col_names = TRUE,
              show_col_types = FALSE) %>%
  drop_na() %>%
  pivot_longer(
    Jan:Dec, 
    names_to = "month", 
    values_to = "unemployment_percent") %>%
  janitor::clean_names() %>%
  arrange(year, month)  
  ncol(unemployment) ` variables. It concludes the key variables such as year, month and unemployment_percent. It shows the percentage of unemployment in each month from Apr/1948 to Sep/2014.
```{r}
library(tidyverse)
unemployment = read_csv("./data/p3/unemployment.csv",
              col_names = TRUE,
              show_col_types = FALSE) %>%
  drop_na() %>%
  pivot_longer(
    Jan:Dec, 
    names_to = "month", 
    values_to = "unemployment_percent") %>%
  janitor::clean_names() %>%
  arrange(year, month)
  
unemployment
```

### Merging 
Join the datasets by merging snp into pols, and merging unemployment into the result.

Overall description:
The separate descriptions are shown before. According to the merged dataset above, we can get the following information. It includes `r pols_snp = left_join(pols, snp, by = c("year", "month"))
pols_snp_unemployment = left_join(pols_snp, unemployment, by = c("year", "month")) 
nrow(pols_snp_unemployment) ` observations and `r pols_snp = left_join(pols, snp, by = c("year", "month"))
pols_snp_unemployment = left_join(pols_snp, unemployment, by = c("year", "month"))  
ncol(pols_snp_unemployment) ` variants. It shows the number of governors/ senators/ representatives and president of republican/ democratic every month from Jan/1947 to Jun/2015. The key variable president shows whether the president is a republican (gop = 1/2) or democratic (dem = 1). It also shows the closing dates and  values of the S&P stock index, and percentage of unemployment every month from Jan/1947 to Jun/2015.
```{r}
pols_snp = left_join(pols, snp, by = c("year", "month"))
pols_snp_unemployment = left_join(pols_snp, unemployment, by = c("year", "month"))
nrow(pols_snp_unemployment) 
ncol(pols_snp_unemployment)

pols_snp_unemployment
```

## Problem 1
### Read and clean
Read and clean the data; retain line, station, name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance. Convert the entry variable from character (YES vs NO) to a logical variable (the ifelse or recode function may be useful).

There are `r nyc_transit = read_csv("./data/p1.csv",
              col_names = TRUE,
              show_col_types = FALSE) %>%
  janitor::clean_names() %>%
  select(-division, -exit_only,-staffing, -staff_hours, -ada_notes, -free_crossover, -north_south_street, -east_west_street, -corner, -entrance_latitude, -entrance_longitude, -station_location, -entrance_location) %>%
  mutate(route8 = as.character(route8), route9 = as.character(route9), route10 = as.character(route10), route11 = as.character(route11)) %>%
  pivot_longer(
    route1:route11, 
    names_to = "route", 
    values_to = "route_value") %>%  
  filter(!is.na(route_value) ) %>%
  unique() %>%
  mutate(entry = recode(entry, YES = TRUE, NO = FALSE)) 
  nrow(nyc_transit)` observations and `r nyc_transit = read_csv("./data/p1.csv",
              col_names = TRUE,
              show_col_types = FALSE) %>%
  janitor::clean_names() %>%
  select(-division, -exit_only,-staffing, -staff_hours, -ada_notes, -free_crossover, -north_south_street, -east_west_street, -corner, -entrance_latitude, -entrance_longitude, -station_location, -entrance_location) %>%
  mutate(route8 = as.character(route8), route9 = as.character(route9), route10 = as.character(route10), route11 = as.character(route11)) %>%
  pivot_longer(
    route1:route11, 
    names_to = "route", 
    values_to = "route_value") %>%  
  filter(!is.na(route_value) ) %>%
  unique() %>%
  mutate(entry = recode(entry, YES = TRUE, NO = FALSE)) 
  ncol(nyc_transit)` variables in the dataset. It concludes key variables such as line, station, name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance.
  
So far, the data is clearer than before but not clear enough. So I use **pivot_longer** to make the routes variables into one column "route" and it is clearer. 
```{r}
library(tidyverse)
nyc_transit = read_csv("./data/p1.csv",
              col_names = TRUE,
              show_col_types = FALSE) %>%
  janitor::clean_names() %>%
  select(-division, -exit_only,-staffing, -staff_hours, -ada_notes, -free_crossover, -north_south_street, -east_west_street, -corner, -entrance_latitude, -entrance_longitude, -station_location, -entrance_location) %>%
  mutate(route8 = as.character(route8), route9 = as.character(route9), route10 = as.character(route10), route11 = as.character(route11)) %>%
  pivot_longer(
    route1:route11, 
    names_to = "route", 
    values_to = "route_value") %>%  
  filter(!is.na(route_value) ) %>%
  unique() %>%
  mutate(entry = recode(entry, YES = TRUE, NO = FALSE))

nyc_transit
         
```

### The number of stations
There are `r nyc_transit %>%
  distinct(station_name, line) %>% 
  nrow()` different stations according to the station_name and line.
```{r}
nyc_transit %>%
  distinct(station_name, line) %>% 
  nrow()
```

### ADA compliant
There are `r nyc_transit %>%
  filter(ada == "TRUE") %>%
  distinct(station_name, line) %>% 
  nrow()` stations are ADA compliant.
```{r}
nyc_transit %>%
  filter(ada == "TRUE") %>%
  distinct(station_name, line) %>% 
  nrow()
```

### Proportion
What proportion of station entrances / exits without vending allow entrance?
The proportion of station entrances / exits without vending allow entrance is 0.434.
```{r}
without_vending = 
  nyc_transit %>%
  filter(vending == "NO") %>%
  distinct(station_name, line)
  nrow(without_vending)

entry_without_vending =
  nyc_transit %>%
  filter(vending == "NO", entry == "TRUE") %>%
  distinct(station_name, line)
  nrow(entry_without_vending)

proportion = nrow(entry_without_vending) / nrow(without_vending) 
proportion
```

### Reformat
Reformat data so that route number and route name are distinct variables. How many distinct stations serve the A train? 

I already tidied the dataset before to make routes into 1 column. Now I will reformat and do it again and there are 60 stations serve the A train.
```{r}
library(tidyverse)
reformat = read_csv("./data/p1.csv",
              col_names = TRUE,
              show_col_types = FALSE) %>%
  janitor::clean_names() %>%
  select(-division, -exit_only,-staffing, -staff_hours, -ada_notes, -free_crossover, -north_south_street, -east_west_street, -corner, -entrance_latitude, -entrance_longitude, -station_location, -entrance_location) %>%
  mutate(route8 = as.character(route8), route9 = as.character(route9), route10 = as.character(route10), route11 = as.character(route11)) %>%
  pivot_longer(
    route1:route11, 
    names_to = "route", 
    values_to = "route_value") %>%  
  filter(!is.na(route_value), 
         route_value == "A") %>%
  distinct(station_name, line) %>%
  nrow()

reformat
```

Of the stations that serve the A train, how many are ADA compliant.
17 of them are ADA compliant.
```{r}
ada_compliant = reformat = read_csv("./data/p1.csv",
              col_names = TRUE,
              show_col_types = FALSE) %>%
  janitor::clean_names() %>%
  select(-division, -exit_only,-staffing, -staff_hours, -ada_notes, -free_crossover, -north_south_street, -east_west_street, -corner, -entrance_latitude, -entrance_longitude, -station_location, -entrance_location) %>%
  mutate(route8 = as.character(route8), route9 = as.character(route9), route10 = as.character(route10), route11 = as.character(route11)) %>%
  pivot_longer(
    route1:route11, 
    names_to = "route", 
    values_to = "route_value") %>%  
  filter(!is.na(route_value), 
         route_value == "A",
         ada == "TRUE") %>%
  distinct(station_name, line) %>%
  nrow()

ada_compliant
```
